WASP-4 is accelerating towards the Earth: A Cautionary Tale in the Search for Tidal Orbital Decay

### 2020-01-23 paper/ms.tex
wrote some words

\documentclass[12pt,twocolumn,tighten]{aastex62}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}

@@

% TODO: may want to improve title
\title{WASP-4 is accelerating towards the Earth}

@@

The main focus of this study is the hot Jupiter WASP-4b, which has an
orbital period that from transits appears to be decreasing by about 10
milliseconds per year.
After discovering the timing variations in the TESS data (CITE Bouma
2019), we obtained N additional radial velocity measurements using
Keck-HIRES, extending the RV baseline by Y years.
Previously, the radial velocities were sparse and were consistent with
not showing any linear trend.
Our new measurements reveal a line of sight acceleration of
$\dot{\gamma} = -XX {\rm m\,s^{-1}\,yr^{-1}}$.
This translates to an expected period decrease simply from the light
travel time effect of ZZ milliseconds per year --- about what is
observed.

% FIXME: figure out correct sections
\S~\ref{sec:observations} presents the new observations...
\S~\ref{sec:timing} describes our timing analysis...
\S~\ref{sec:implications} ...
\S~\ref{sec:future}...

\section{WASP-4's acceleration towards the Earth}
\label{sec:observations}

\subsection{Trend due to stellar activity?}

First, we considered whether the measured RV trend is correlated with
WASP-4's emission in the Ca II H \& K lines, as quantified with the
chromospheric $S$-index (CITE: Wright 2004).  We only analyzed the
HIRES velocities for this step, since they are the main source of
signal in our analysis.  First, we subtracted the orbital solution
from the Keck-HIRES velocities.  Then, following CITE Bryan 2016 and
2019, we calculated the Spearman rank correlation coefficient between
the $S$-index and the orbit-subtracted velocities.  We found a
correlation coefficient of 0.16, which weakly hints at a correlation,
but is not statistically significant (the corresponding $p$-value is
0.65).  We therefore do not think that the linear trend can be
attributed to stellar activity.

\subsection{Zorro imaging}

Convinced that the linear trend warranted closer examination, we
performed speckle imaging using Zorro at Gemini to attempt to detect
any nearby companions.  To our knowledge, this is the first high
resolution imaging that has been performed for WASP-4 (FIXME: IS THIS
TRUE?).

We observed WASP-4 twice with Gemini-South, 11/12 Sept. (8 sets of
1000 frames each) and 28/29 Sept. (7 sets of 1000 frames each). The
second set had better seeing and produced a slightly better result.
The spatial scale is from the diffraction limit of Gemini at the
wavelength observed to about 1.25'' plus a bit more.

% TODO: might Howell suggest a paragraph for how the contrast ratios
% were derived?
To derive contrast ratios from these images, we did XXXXX.

We would like to convert the contrast ratios obtained through the
Zorro imaging to constraints on the masses of putative companions, and
their separations from the host star.

To do this, we followed the methodology of CITE Montet+2014, and opted
to employ the CITE Baraffe+2003 models for substellar mass objects and
the MESA isochrones for stellar mass objects (CITE Choi et al XXXX).
In the intermediate regime, near 100 Mjup, we interpolated between the
two sets of models.

We assumed that companions would have contracted, and asserted the
system age to be 5 Gyr.

Due to the narrow-band filters of the Zorro imager, we further assumed
that all sources had blackbody spectra. While this is clearly false,
we do not readily have access to the planetary and stellar atmosphere
models needed for the correct calculation. However, the CITE
Baraffe+2003 and MESA models do report effective temperatures and
bolometric luminosities.  We opt to use these theoretical quantities,
combined with the measured Zorro filter wheels, to the absolute
magnitudes in the 562 and 832 nm Zorro bandpasses.  The apparent
magnitudes follow from information about the distance of the star.

Applying the same assumption to WASP-4 itself, we compute its apparent
magnitude in each of the Zorro filters, and finally derive the
transformation from contrast to companion mass.

Given a linear radial velocity trend, we can place limits on the mass
and semi-major axis of any planetary or stellar companions in the
system.
The constraints on the two parameters are degenerate: $\dot{\gamma}
\propto M_{\rm p} a^{-2}$.
Nonetheless, the additional information from the Zorro images can help
limit the available parameter space by setting an outer limit on the
semimajor axis, and a maximum brightness (and thereby mass) of any
putative companions.

The general procedure we use to derive these constraints has been developed by
\citet{wright_linear_trends_2007}, \citet{montet_trends_2014},
\citet{knutson_friends_2014}, \citet{bryan_statistics_2016,bryan_excess_2019},
and others.
For this study, we closely follow the procedure of \citet{bryan_excess_2019}.

We begin by defining a $50\times50$ grid in true planetary mass and semimajor
axis, evenly spaced from 1 to 300$\,$$M_{\rm Jup}$ and 3 to 500$\,$${\rm AU}$.

We then consider the possibility that an additional companion in any particular
cell can explain the observed linear trend.

In each cell, we consider 500 hypothetical companions.

We assign each companion a mass and semimajor axis from log-uniform
distributions within the grid cell. We draw the inclination from a uniform
distribution in $\cos i$.  The eccentricity is drawn from
\citet{kipping_beta_2013}'s long-period exoplanet Beta distribution ($a=1.12$,
$b=3.09$).

For each simulated companion, we then draw a sample from the converged chains
of our initial model of WASP-4b, plus its linear trend. We subtract
the planet's orbital solution, leaving the linear trend.

Given $(a_{\rm c}, M_{\rm c}, e_{\rm c}$ for each simulated outer companion,
and the fixed instrument offsets and jitters from the MCMC chains, we then
perform a maximum likelihood fit for the time and argument of periastron of the
outer simulated companion.

We then convert the resulting $50\times50\times500$ cube of best-fit
log-likelihood values to probabilities, and average over the samples in
each grid cell to derive a probability distribution in mass and semi-major
axis.

The distribution is shown in {\bf Figure~X}.

\section{Do other hot Jupiters show the Romer delay?}
yes, HAT-P-7b. but omg utc to tdb.

\section{What is the expected Pdot distribution?}

SUMMARY OF LITERATURE

This is a major challenge in the hunt for orbital decay.  The main
example of this study is the hot Jupiter WASP-4b, which we recently
showed transited $\approx$82 seconds early for the TESS spacecraft
assuming a constant orbital period \citep{bouma_wasp-4b_2019}.  Our
timing analysis included data from peer-reviewed literature for which
the times were measured from a single transit, and for which the
midpoint was allowed to be a free parameter. We also required the time
system to be clearly documented\footnote{In other words, we needed to
know whether any heliocentric or barycentric corrections had been
performed, and whether the absolute time system was UTC or TDB.}. The
transit times we analyzed spanned 2007 to 2019
\citep{wilson_wasp-4b_2008,gillon_improved_2009,winn_transit_2009,dragomir_terms_2011,sanchis-ojeda_starspots_2011,nikolov_wasp-4b_2012,hoyer_tramos_2013,ranjan_atmospheric_2014,huitson_gemini_2017}.
The combined timing data were best fit by an ephemeris with a constant
negative period derivative; our best-fit decay rate was $\dot{P} =
-12.6 \pm 1.2$ milliseconds per year.  Our interpretation was that the
apparent period change could be caused by any of three scenarios: a
decaying orbit, a precessing orbit, and an orbit being gravitationally
perturbed by an outer companion.

@@

One line of follow-up needed to confirm and understand the timing
variation was additional radial velocity observations.  The longest RV
baseline previously available was six observations acquired by
\citet{knutson_friends_2014} and the CKS team on Keck HIRES from
2010 to 2013.  This past season we acquired four additional
observations with Keck HIRES. Even before fitting out the hot Jupiter,
the residuals showed a strong linear trend.

For completeness, in our analysis we included CORALIE measurements
from \citet{wilson_wasp-4b_2008} and \citet{triaud_spin-orbit_2010},
using the homogeneous velocities reported by the latter authors.  We
included the HARPS values reported by \citet{pont_determining_2011},
which are identical to those from \citet{husnoo_observational_2012}.
We omitted the HARPS data points taken over three nights by
\citet{triaud_spin-orbit_2010} for Rossiter-McLaughlin observations
because they were calculated using a different pipeline than the
longer-baseline \citeauthor{pont_determining_2011} measurements, and
necessarily inclusion of an extra offset term would nullify their
statistical value.

We then fitted a single Keplerian orbit, plus instrument offsets,
jitters, and a linear trend
\citep[][\texttt{radvel}]{fulton_radvel_2018}.  We set Gaussian priors
on the period and time of inferior conjunction using the values from
\citet{bouma_wasp-4b_2019}, and fixed the eccentricity to zero,
consistent with results from \citet{beerer_secondary_2011},
\citet{knutson_friends_2014} and \citet{bonomo_gaps_2017}.  The
remaining free parameters were the velocity semi-amplitude, the
instrument zero-points, the instrument jitters (an additive white
noise term for each instrument), and optionally a linear acceleration
term ($\dot{v_{\rm r}}$).

%FoHJ: gammadot = -0.0099^{+0.0052}_{-0.0054}
The AIC and BIC strongly favored a model with a linear trend ($\Delta
{\rm BIC}=73$ compared to a model without the linear trend).
Figure~\ref{fig:rv_o_minus_c} shows the best-fitting model, with the
orbit of WASP-4b subtracted.  The best-fit radial velocity derivative,
$\dot{v}_{\rm r}$, is
\begin{equation}
  \dot{v}_{\rm r} = -0.0422^{+0.0028}_{-0.0027}\,{\rm m\,s^{-1}\,day^{-1}}.
\end{equation}
This acceleration towards our line of sight is 4.3 times faster than
the $\approx$$2\sigma$ trend \citet{knutson_friends_2014}
found from the shorter baseline.  The radial velocity values are
available in the data-behind-the-figure online version of this {\it
Note}.

Under the assumption of constant acceleration, $\dot{P} = \dot{v}_{\rm
r} P / c$, the implied period decrease that should be seen in
transits is roughly $5.9 \pm 0.4$ milliseconds per year.  All the
available TTV data show a period decrease of $\approx 8-12$
milliseconds per year~\citep{bouma_wasp-4b_2019,southworth_transit_2019,baluev_homogeneously_2019}.

Though the quantitative agreement is not perfect, Occam's razor would
suggest that the line of sight acceleration is probably a sufficient
explanation for the apparent decrease of WASP-4b's orbital period.
While further radial velocity observations of the system should help
in determining the mass and semi-major axis of the companion, it seems
unlikely that additional transit observations will yield near-term
constraints on orbital decay.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% RNAAS: you get one figure.

\begin{figure}
    \begin{center}
		\includegraphics{rv_fit.pdf}
    \end{center}
    \vspace{-0.8cm}
    \caption{
      {\bf Radial velocity observations of WASP-4.}
      The orbit of WASP-4b has been subtracted.  The best-fit linear
      trend from the RVs is shown with the black line, with $1\sigma$
      errors in gray. The trend needed to produce the period decrease
      of 12.6 milliseconds per year seen by \citet{bouma_wasp-4b_2019}
      in transits is the purple dotted line in the bottom left.
     \label{fig:rv_o_minus_c}
    }
\end{figure}

@@

\acknowledgements
%
This paper includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
This work made use of NASA's Astrophysics Data System Bibliographic
Services.
%
This research has made use of the VizieR catalogue access tool, CDS,
Strasbourg, France. The original description of the VizieR service was
published in A\&AS 143, 23.
%
This work has made use of data from the European Space Agency (ESA)
mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
for the DPAC has been provided by national institutions, in particular
the institutions participating in the {\it Gaia} Multilateral
Agreement.
%
\newline
%
\software{
	\texttt{astroplan} \citep{astroplan2018},
  \texttt{radvel} \citep{fulton_radvel_2018}
  %FIXME numerical python packages
}

>>>>

\documentclass[12pt,twocolumn,tighten]{aastex62}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{natbib}

@@

\defcitealias{bouma_wasp4b_2019}{B19}

% TODO: may want to improve title
\title{WASP-4 is accelerating towards the Earth: A Cautionary Tale in
the Search for Tidal Orbital Decay}
% "An Explanation for WASP-4's Decreasing Orbital Period".
% "WASP-4's Decreasing Orbital Period: Solved."

@@

The main focus of this study is the hot Jupiter WASP-4b, which has an
orbital period that from transits appears to be decreasing by about 10
milliseconds per year.
After discovering the timing variations in the TESS data 
\citep[][hereafter \citetalias{bouma_wasp4b_2019}]{bouma_wasp4b_2019}, we obtained N additional
radial velocity measurements using
Keck-HIRES, extending the RV baseline by Y years.
Previously, the radial velocities were sparse and were consistent with
not showing any linear trend.
Our new measurements reveal a line of sight acceleration of
$\dot{\gamma} = -XX {\rm m\,s^{-1}\,yr^{-1}}$.
This translates to an expected period decrease simply from the light
travel time effect of ZZ milliseconds per year --- about what is
observed.

@@

% FIXME: figure out correct sections
\S~\ref{sec:observations} presents the new observations...
\S~\ref{sec:analysis} describes our analysis...
\S~\ref{sec:discussion} discusses...
\S~\ref{sec:conclusion} gives conclusions...

\section{Observations}
\label{sec:observations}

\subsection{Transits}

{\bf Table~N} lists the transit times we used in our analysis.  We
include data from the peer-reviewed literature for which (a) the
analysis was based on observations of a single transit, (b) the
midpoint was fitted as a free parameter, and (c) the time system
specified both the absolute reference (TDB or UTC) and also whether
any barycentric or heliocentric corrections had been performed.

The majority of times are identical to those we collected in
\citetalias{bouma_wasp4b_2019}.  Twenty-two new times reported by
\citet{southworth_transit_2019} are included.  These transits were
observed from the 3.58m NTT and Danish 1.54m telescopes at La Silla,
and the SAAO 1.0m telescope.

% FIXME : verify with Michael Gillon and Phil Evans that this is true.
A number of additional times were also recently reported by
\citet{baluev_2019}, based on a homogeneous analysis of archival
ground-based observations.  We included their ``high quality'' transit
times with provenance from the TRAPPIST and El Sauce telescopes, for
which we verified with the original observers that appropriate
time-system corrections had been performed (M.~Gillon, P.~Evans,
priv{.} comm{.}).

The four available occultation data points tabulated by
\citetalias{bouma_wasp4b_2019} are sparse and have neglegible
statistical value due to their large uncertainties, and we forgo
their use in this analysis.

\subsection{Radial velocities}

% TODO: Details of reduction from Howard or Knutson
After identifying the period decrease in
\citetalias{bouma_wasp4b_2019}, we acquired four additional radial
velocity measurements with the Keck High Resolution Echelle
Spectrometer (HIRES, \citealt{vogt_hires_1994}).  Our observations
were acquired under the purview of the California Planet Survey
\citep{howard_cps_2010}.  The spectra were reduced using {\bf software
X}.  Previously, the HIRES data-points spanned 2010 to 2013
\citep{knutson_friends_2014}.  Our new measurements triple the HIRES
observing baseline to nine years.

%FIXME: I should ask Didier if he has any new CORALIE observations...
The complete set of radial velocity observations is given in {\bf
Table~M}.  Along with the 2010-2019 HIRES observations, there are also
earlier measurements from CORALIE and HARPS.  Following
\citetalias{bouma_wasp4b_2019}, we included the CORALIE measurements
from \citet{wilson_wasp-4b_2008} and \citet{triaud_spin-orbit_2010},
using the homogeneous radial velocities calculated by the latter
authors. We included the HARPS values reported by
\citet{pont_determining_2011} and \citet{husnoo_observational_2012}.
We omitted the HARPS data points taken over three nights by
\citet{triaud_spin-orbit_2010} for Rossiter-McLaughlin observations
because they have a systematic offset from the remaining datasets.

\subsection{Speckle imaging}

A cursory analysis of the new HIRES observations led to our detection
of a linear trend in the residuals after fitting out the orbit of
WASP-4b.  This prompted us to acquire speckle imaging using Zorro at
Gemini-South \citep[see][, and the instrument
web-pages\footnote{\url{www.gemini.edu/sciops/instruments/alopeke-zorro/}}]{scott_nessi_2018}.
Zorro (and its counterpart, 'Alopeke, on Gemini-North) are
dual-channel speckle interferometers that can be used to observe in
narrow-band filters centered at 562$\,$nm and 832$\,$nm.

We observed WASP-4 twice, on the night of September 11-12 (8 sets of
1000 frames each) and also on the night of September 28-29 (7 sets of
1000 frames each). The second night had better seeing and produced a
slightly better result, which we opt to use for our analysis.

% TODO: might Howell suggest a paragraph for how the contrast ratios
% were derived?
{\bf We reduced the speckle images to contrast curves by ...}

\section{Analysis}
\label{sec:analysis}

\subsection{Transits}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Timing residuals and best-fit models for WASP-4b.}
  The vertical axis shows the observed times minus the calculated
  times assuming a constant orbital period .  Darker points correspond
  to more precise data.  The quadratic ephemeris and its $1\sigma$
  uncertainties are shown in blue.  The binned TESS point (red square)
  is the weighted average of 18 TESS transits and is for display
  purposes only.  The models were fitted to all of the individual
  transit times.
  \label{fig:times}
	}
\end{figure*}

We considered two possible models for the observed transit times.
The first model assumes a constant orbital period on a circular orbit:
\begin{align}
  t_{\rm tra}(E) &= t_0 + PE,\\
\end{align}
where $E$ is the epoch number and $t_0$ is a reference epoch.
The second model assumes the period changes at a steady rate:
\begin{align}
  t_{\rm tra}(E) &=
    t_0 + PE +
    \frac{1}{2} \frac{{\rm d}P}{{\rm d}E} E^2.
\end{align}
The free parameters are the reference epoch $t_0$, the period at
the reference epoch, and the period derivative, ${\rm d}P/{\rm d}t =
(1/P) {\rm d}P/{\rm d}E$.
We defined the epoch numbers such that $E=0$ is near the weighted
average of the observed times.  This helps to reduce the covariance
between $t_0$ and $P$.
A third possible model that we did not consider for reasons that will
become apparent is a precessing, eccentric orbit \citep[{\it
e.g.},][]{gimenez_revision_1995,patra_2017}.

We fitted each of the two models by assuming a Gaussian likelihood and
sampling over the posterior probability distributions.  We sampled the
posterior using the algorithm proposed by
\citet{goodman_ensemble_2010} and implemented by
\citet{foreman-mackey_emcee_2013} in \texttt{emcee}.  The prior for
the quadratic model allowed the period derivative to have any sign.

Figure~\ref{fig:times} shows the observed transit times, minus the
best-fit constant period model. 
The best-fitting constant-period model has $\chi^2 = XXX$ and YY
degrees of freedom.  The best-fitting
quadratic model has $\chi^2 = YYY$ and ZZ degrees of freedom. 
The difference in the BIC \citep{kass_bayes_1995} between the linear
and quadratic and models is $\Delta {\rm BIC} = ZZ.Z$, strongly
favoring the quadratic case.

The best-fit period derivative in the quadratic model is
\begin{equation}
\dot{P}
  = - (X.XX \pm 0.XX)\times 10^{-10}
  = - YY.Y \pm Z.Z~{\rm ms}\,{\rm yr}^{-1}.
  \label{eq:dP_dt_obs}
\end{equation}
This agrees with the value reported by \citet{southworth_transit_2019}
after their including an additional 22 archival transit measurements
($\dot{P} = - 9.2 \pm 1.1~{\rm ms}\,{\rm yr}^{-1}$).
It is smaller than the rate of period decrease found by
\citetalias{bouma_wasp4b_2019} ($- 12.6 \pm 1.2~{\rm ms}\,{\rm
yr}^{-1}$)., presumably because of missing coverage filled in by
\citeauthor{southworth_transit_2019}'s observations.
The other best-fit transit timing model parameters are reported in
{\bf Table~X}.

\subsection{Radial Velocities: WASP-4's acceleration towards the Earth}

%FIXME: is the purple here from my Pdot, or Southworth's?
%FIXME: perhaps include the WASP-4b phase-fold here too??
\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.48\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Radial velocity observations of WASP-4.} The best-fit Keplerian orbit
    of WASP-4b has been subtracted.  The linear trend inferred from the
    RV data is shown with a black line, with $1\sigma$ errors in gray.
    The trend that would be needed to produce the period decrease seen
    in transits ($\dot{P}= - YY.Y \pm Z.Z~{\rm ms}\,{\rm yr}^{-1}$) is
    indicated with the purple dotted line.  The four new RV
    measurements presented in this work increased the significance of
    the slope from $\approx$2$\sigma$ to 15$\sigma$.
	\label{fig:rvs}
  \vspace{-0.3cm}
	}
\end{figure}

%FIXME: this entire paragraph copies Bouma+19 too heavily.
We began by fitting a single Keplerian orbit, plus instrument offsets,
jitters, and a long-term trend
\citep[][\texttt{radvel}]{fulton_radvel_2018}.  We set Gaussian priors
on the period and time of inferior conjunction using the values from
Table~4 of \citetalias{bouma_wasp4b_2019}, and fixed the eccentricity
to zero, per the results of \citet{beerer_secondary_2011},
\citet{knutson_friends_2014} and \citet{bonomo_gaps_2017}.  The free
parameters were the velocity semi-amplitude, the instrument
zero-points, the instrument jitters (an additive white noise term for
each instrument), linear ($\dot{v_{\rm r}}$), and optionally
second-order ($\ddot{v_{\rm r}}$) acceleration terms.

%FIXME: this is a janky approach if so??
We found that the best-fitting model with both linear and quadratic
radial velocity terms was marginally preferred (by $\Delta
\mathrm{BIC} = 5.8$) over the best-fitting model with only a linear
term.  Regardless, for consistency with \citet{knutson_friends_2014},
who fixed the quadratic component of the long-term trend to zero, in
Figure~\ref{fig:rvs} we show best-fitting models for the linear-trend
case.

WASP-4 was found to be accelerating towards our line of sight at high
confidence,
\begin{equation}
  \dot{v}_{\rm r} =
     -0.0422^{+0.0028}_{-0.0027}
     \,{\rm m}\,{\rm s}^{-1}\,{\rm day}^{-1}.
\end{equation}
For comparison, before our recent observing run, 
$\dot{v}_{\rm r}$ was thought to be about five times smaller, and was
only marginally significant
\citep{knutson_friends_2014,bouma_wasp4b_2019}.

The system's acceleration towards our line of sight causes a decrease
in the apparent orbital period (the R{\o}mer delay).  The period
derivative expected from radial velocities is
\begin{align}
  \dot{P}_{\rm RV} &= \frac{\dot{v}_{\rm r} P}{c} \\
  \dot{P}_{\rm RV} &= -5.94 \pm 0.39~{\rm ms}\,{\rm yr}^{-1}.
\end{align}
In other words, the majority of the period decrease observed in
transits ($\dot{P} = - X.X \pm 1.1~{\rm ms}\,{\rm yr}^{-1}$) seems to
be explained by the acceleration of the host star.
% FIXME
All the
available TTV data show a period decrease of $\approx 8-12$
milliseconds per year~\citep{bouma_wasp-4b_2019,southworth_transit_2019,baluev_homogeneously_2019}.

An important consideration is whether the measured RV trend is at all
correlated with stellar activity.  We investigate this by analyzing
WASP-4's emission in the Ca II H \& K lines, as quantified with the
chromospheric $S$-index \citep{wright_chromospheric_2004}.  We only
examined the HIRES velocities for this step, since they are the main
source of signal for our analysis.  First, we subtracted the orbital
solution from the Keck-HIRES velocities.  Then, following
\citet{bryan_statistics_2016,bryan_excess_2019}, we calculated the
Spearman rank correlation coefficient between the $S$-index and the
orbit-subtracted velocities.  We found a correlation coefficient of
0.16. Though suggestive, it is not statistically significant (the
corresponding $p$-value is 0.65).  Furthermore, inspection of the
$S$-index timeseries showed no secular or sinusoidal trends, as would
be expected if we were observing a long-term magnetic activity cycle.
The $S$-index values are included in {\bf Table X}.  We conclude that
it would be highly unlikely for the linear trend to be caused by
stellar activity.

%TODO: describe 3b better
\begin{figure*}[!t]
	\gridline{
		\fig{f3a.png}{0.33\textwidth}{}
		\fig{f3b.pdf} {0.5\textwidth}{}
	}
	\vspace{-0.7cm}
    \caption{
      {\it Left:\,} 
      Speckle image reconstructed from 1000 40 millisecond frames in an 832 nm bandpass, and acquired on September 28, 2019.
      The image scale is $2.46''\times2.46''$.
      {\it Right:\,} 
      Contrast curve derived from point-source injection-recovery
      experiments.
    }
    \label{fig:zorro}
\end{figure*}

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.48\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Masses and semi-major axes of companions that meet
    requirements of both the radial velocity and the speckle imaging.}
    The likelihood inferred from radial velocities is shown in
    grayscale, and the region excluded from the imaging is shown with
    a cross-hatch pattern.
	\label{fig:mass_sma}
  \vspace{-0.3cm}
	}
\end{figure}

Given a linear radial velocity trend, we can place lower-limits on the
mass and semi-major axis of additional bodies in the system.  The
constraints on the two parameters are degenerate: $\dot{\gamma}
\propto M_{\rm p} a^{-2}$.  Nonetheless, the Zorro images and contrast
curves can limit the available parameter space by setting an upper
limit on the semimajor axis, and a maximum brightness (and thereby
mass) of any putative companions.

The general procedure we use to derive constraints on possible
companion masses and semi-major axes has been developed by
\citet{wright_linear_trends_2007}, \citet{crepp_trends_2012},
\citet{montet_trends_2014}, \citet{knutson_friends_2014},
\citet{bryan_statistics_2016,bryan_excess_2019}, and others.

\paragraph{Speckle imaging constraints}

First, we would like to convert the contrast ratios obtained through
the Zorro imaging (Figure~\ref{fig:zorro}) to limits on the masses of
putative companions and their separations from the host star.

To do this, we followed \citet{montet_trends_2014}, and opted to
employ the \citet{baraffe_evolutionary_2003} models for substellar
mass objects and the MIST isochrones for stellar mass objects
\citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015,dotter_mesa_2016,choi_mesa_2016}.
We assumed that the system age was 5 Gyr, so that companions would
have fully contracted.

Due to the custom filters of the Zorro imager, and corresponding lack
of synthetic photometry, we further assumed that all sources had
blackbody spectra. While this is clearly false, we do not readily have
access to the planetary and stellar atmosphere models needed for the
consistent calculation with the \texttt{COND03} and \texttt{MESA}
models.  We therefore adopted the effective temperatures and
bolometric luminosities from the \citet{baraffe_evolutionary_2003} and
MIST isochrones.  Using these theoretical quantities and the empircal
Zorro bandpass functions, we calculated absolute magnitudes in the 562
and 832 nm Zorro bandpasses for stellar and planetary mass companions.
Applying the same calculation to WASP-4 itself using the effective
temperature and bolometric luminosity from
\citetalias{bouma_wasp4b_2019}, we derived the transformation from
contrast ratio to companion mass.
The resulting limits are shown as the cross-hatched region in
Figure~\ref{fig:mass_sma}.

\paragraph{Radial velocity constraints}

To derive constraints on possible companion masses and separations
from the radial velocities, we followed the procedure of
\citet{bryan_excess_2019}.  We began by defining a $50\times50$ grid
in true planetary mass and semimajor axis, evenly spaced from 1 to
300$\,$$M_{\rm Jup}$ and 3 to 500$\,$${\rm AU}$.  We then considered the
possibility that an additional companion in any particular cell could
explain the observed linear trend.  In each cell, we simulated 500
hypothetical companions.

We assigned each companion a mass and semimajor axis from log-uniform
distributions within the grid cell. We drew the inclination from a
uniform distribution in $\cos i$.  The eccentricity was drawn from
\citet{kipping_beta_2013}'s long-period exoplanet Beta distribution
($a=1.12$, $b=3.09$).  For each simulated companion, we then drew a
sample from the converged chains of our initial model of WASP-4b, plus
its linear trend. We subtracted the planet's orbital solution, leaving
the linear trend.  Given $(a_{\rm c}, M_{\rm c}, e_{\rm c}$ for each
simulated outer companion, and the fixed instrument offsets and
jitters from the MCMC chains, we then performed a maximum likelihood
fit for the time and argument of periastron of the outer simulated
companion.  We converted the resulting $50\times50\times500$ cube of
best-fit log-likelihood values to probabilities, and averaged over the
samples in each grid cell to derive a probability distribution in mass
and semi-major axis.

The result is shown as grayscale background in
Figure~\ref{fig:mass_sma}.

\section{Discussion}
\label{sec:discussion}

\subsection{Implications for WASP-4}
Previous possibilities suggested to explain the period decrease of
WASP-4b included tidal orbital decay, orbital precession, and
light-travel time effects \citep{bouma_wasp4b_2019}.  Our new radial
velocity measurements have now shown the least exotic
option---light-travel time effects---is also the most likely
explanation.  Transits show the obital period to decrease by $YY.Y \pm
Z.Z~{\rm ms}\,{\rm yr}^{-1}$; the line-of-sight acceleration observed
in radial velocities implies an expected period decrease of $YY.Y \pm
Z.Z~{\rm ms}\,{\rm yr}^{-1}$.  
Though the quantitative agreement is not perfect, Occam's razor would
suggest that the line of sight acceleration is probably a sufficient
explanation for the apparent decrease of WASP-4b's orbital period.
% While further radial velocity observations of the system should help
% in determining the mass and semi-major axis of the companion, it seems
% unlikely that additional transit observations will yield near-term
% constraints on orbital decay.

The corresponding requirements for the companion causing the
acceleration are that it is likely either a brown-dwarf or low mass
star, and that it is between $10$-$100\,{\rm AU}$ from the host star
(Figure~\ref{fig:mass_sma}).
Given such a mass, this companion could at one time have influenced
the orbital evolution of the inner giant.
Further radial velocity monitoring should eventually reveal the
orbital parameters and minimum mass of the companion.

\subsection{How many other hot Jupiters will show the R\'omer delay?}

We identified WASP-4b's decreasing orbital period as part of a search
for tidal orbital decay.  However, over half of hot Jupiters have
companions outside of $1\,{\rm AU}$ with super-Jovian masses
\citep{knutson_friends_2014}.  Line-of-sight accelerations are
therefore relatively common in hot Jupiter systems.

To evaluate the importance of these effects in future transit timing
searches, we collected the linear trends reported by
\citet{knutson_friends_2014}, and computed the expected orbital period
changes $\dot{P}_{\rm RV} = \dot{v}_{\rm r} P/c$ for all the
systems.
The results are given in {\bf Table~X}, and visualized for hot
Jupiters with significant linear trends in Figure~\ref{fig:pdot_pop}.

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.48\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf ...}
	\label{fig:pdot_pop}
  \vspace{-0.3cm}
	}
\end{figure}

\section{Conclusions}
\label{sec:conclusions}

@@

\acknowledgements
%
This paper includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
This work made use of NASA's Astrophysics Data System Bibliographic
Services.
%
Based on observations obtained at the Gemini Observatory, which is
operated by the Association of Universities for Research in Astronomy,
Inc., under a cooperative agreement with the NSF on behalf of the
Gemini partnership: the National Science Foundation (United States),
National Research Council (Canada), CONICYT (Chile), Ministerio de
Ciencia, Tecnolog\'{i}a e Innovaci\'{o}n Productiva (Argentina),
Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e Inova\c{c}\~{a}o (Brazil),
and Korea Astronomy and Space Science Institute (Republic of Korea).
%
Observations in the paper made use of the High-Resolution Imaging
instrument Zorro at Gemini-South. Zorro was funded by the NASA
Exoplanet Exploration Program and built at the NASA Ames Research
Center by Steve B. Howell, Nic Scott, Elliott P. Horch, and Emmett
Quigley.
%
This research has made use of the VizieR catalogue access tool, CDS,
Strasbourg, France. The original description of the VizieR service was
published in A\&AS 143, 23.
%
This work has made use of data from the European Space Agency (ESA)
mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
for the DPAC has been provided by national institutions, in particular
the institutions participating in the {\it Gaia} Multilateral
Agreement.
%
\newline
%
\software{
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
	\texttt{astroplan} \citep{astroplan2018},
  \texttt{astropy} \citep{the_astropy_collaboration_astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  % \texttt{BATMAN} \citep{kreidberg_batman_2015},
  \texttt{corner} \citep{corner_2016},
  \texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{IPython} \citep{perez_2007},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
	\texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  \texttt{radvel} \citep{fulton_radvel_2018},
  % \texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001}.
}

%\clearpage
%\newpage
\input{model_fit_table.tex}
